---
title: "Valsamo Melanoma SpliceMutr Analysis"
output: html_document
date: "2023-04-25"
---

# Loading libraries

```{r}

library(readxl)
library(ggplot2)
library(DT)
library(dplyr)
library(stringr)
library(rjson)
library(ComplexHeatmap)
library(ggpubr)
library(ggrepel)
library(ggprism)
library(rstatix)
library(ggsignif)
library(effectsize)
library(gridExtra)
library(estimate)

```

# Internal Functions

```{r}

remove_factors <- function(d_frame,g_vector){
  d_frame_filt <- d_frame
  a<-vapply(g_vector,function(g_elem){
    d_frame_filt[,g_elem]<<-unname(unlist(lapply(d_frame[,g_elem],as.character)))
    return(T)
  },logical(1))
  return(d_frame_filt)
}

calculate_cohens_d <- function(p_vals,grouping_variables,comparison_term,input_data){
  data_term <- unique(p_vals$.y.)
  cohens_ds <- t(vapply(seq(nrow(p_vals)),function(row_val){
    curr_group_vars <- p_vals[row_val,grouping_variables,drop=T]
    groups <- c(p_vals[row_val,"group1",drop=T],p_vals[row_val,"group2",drop=T])
    input_data_filt <- input_data
    if (length(grouping_variables)>0){
      for (i in seq(length(grouping_variables))){
        input_data_filt <- input_data_filt[input_data_filt[,grouping_variables[i],drop=T]==curr_group_vars[i],]
      }
    }
    input_data_filt <- input_data_filt %>% dplyr::filter(input_data_filt[,comparison_term,drop=T] %in% groups)
    as.numeric(cohens_d(as.numeric(input_data_filt[,data_term,drop=T]) ~ input_data_filt[,comparison_term,drop=T],data=input_data_filt,iterations=1000))

  },numeric(4)))
  colnames(cohens_ds) <- c("cohens_d","CD_CI","CD_CI_low","CD_CI_high")
  return(data.frame(cohens_ds))
}

```

# Exploring the sample metadata via the study manifest

## Removing faulty samples from manifest and formatting manifest TRTGRP

```{r}

manifest <- read_excel("./input_data/manifest.xlsx")
files_to_remove <- c("hg19MTERCC-ensembl75-genes-Q21777-Plate-1-E06_L65",
"hg19MTERCC-ensembl75-genes-Q21777-Plate-1-F12_L1.D707_508",
"hg19MTERCC-ensembl75-genes-Q23152+B4+H2+AG710464_L1.D705")
manifest <- manifest %>% dplyr::filter(!(Sample %in% files_to_remove))
manifest$Sample <- str_replace_all(manifest$Sample,"hg19MTERCC-ensembl75-genes-","")
fastq_files <- read.table("./input_data/fastq_files.txt")
SJ_files <- read.table("./input_data/SJ_files.txt")
SJ_files$sample_name <- vapply(SJ_files$V1,function(file){
  str_remove(file,"SJ.out.tab")
},character(1))
fastq_files$sample_name <- vapply(fastq_files$V1,function(file){
  str_remove(file,"_1.clipped.fastq.gz")
},character(1))
manifest$AX_TRTGRP <- vapply(manifest$AX_TRTGRP,function(TRTGRP){
  if (str_detect(TRTGRP,"IPI")){
    return("NIV-IPI")
  } else {
    return(TRTGRP)
  }
},character(1))

```

## The number of baseline patients with and without previous exposure to Ipilimumab

```{r}

manifest_table <- data.frame(table(manifest$AX_PriorIPI[manifest$AX_TIMETEMP=="PRE" & manifest$AX_BOR3!="NE"]))
manifest_table$Var1 <- c("No previous IPI","Previous IPI")
datatable(manifest_table)

```

# ESTIMATE Calculations

```{r}

featurecounts_file <- "./input_data/featurecounts_all_vst.rds"
output_file<-"./intermediates/featurecounts_all_vst.gct"
output_estimate_file <- "./intermediates/featurecounts_all_vst.estimate"
GTF_file <- "./input_data/Homo_sapiens.GRCh38.99.rds"

featurecounts_data <- readRDS(featurecounts_file)
GTF <- readRDS(GTF_file)
GTF_filt <- GTF %>% dplyr::filter(V3=="gene")
GTF_filt$gene_name <- vapply(GTF_filt$V9,function(val){
  name_pre <- str_split(val,"; ")[[1]][3]
  name_next<-str_remove(name_pre,"gene_name ")
  gene_name <- str_remove_all(name_next,'"')
  return(gene_name)
},character(1))
GTF_filt$gene_id <- vapply(GTF_filt$V9,function(val){
  name_pre <- str_split(val,"; ")[[1]][1]
  gene_id<-str_remove(name_pre,"gene_id ")
  return(gene_id)
},character(1))
GTF_filt <- GTF_filt[,c("gene_name","gene_id")]

featurecounts_data$ENSEMBL <- vapply(featurecounts_data$ENSEMBL,function(gene){
  a<-which(GTF_filt$gene_id==gene)
  if (length(a)>0 & length(a)<2){
    return(GTF_filt$gene_name[a])
  } else {
    return(NA)
  }
},character(1))
a<-data.frame(table(featurecounts_data$ENSEMBL))
featurecounts_data_filt <- featurecounts_data[!(featurecounts_data$ENSEMBL %in% as.character(a$Var1[a$Freq==2])),]

write.table(featurecounts_data_filt,
            file="./intermediates/featurecounts_all_vst.estimate.txt",quote=F, row.names=F, sep="\t")
featurecounts_file <- "./intermediates/featurecounts_all_vst.estimate.txt"

outputGCT(featurecounts_file,output_file)

GCT_file <- read.table(output_file,sep="\t")
estimateScore(output_file,output_estimate_file,platform="illumina")

ESTIMATE_scores <- t(read.table(output_estimate_file,sep="\t",skip=2,header=T,check.names=F))
colnames(ESTIMATE_scores)<-ESTIMATE_scores["NAME",]
ESTIMATE_scores<-data.frame(ESTIMATE_scores[seq(3,nrow(ESTIMATE_scores)),])
max_estimate <- max(as.numeric(ESTIMATE_scores$ESTIMATEScore))
ESTIMATE_scores$pseudo_purity <- 1 - (as.numeric(ESTIMATE_scores$ESTIMATEScore))/(max_estimate+(max_estimate*0.1))
ESTIMATE_scores$tumor_purity <- cos(0.6049872018+(0.0001467884*as.numeric(ESTIMATE_scores$ESTIMATEScore)))
ESTIMATE_scores$sample <- colnames(featurecounts_data)[seq(2,ncol(featurecounts_data))]
rownames(ESTIMATE_scores)<-ESTIMATE_scores$sample

```

# Evaluating the splicing antigenicity PRE treatment

```{r}

figure_dir <- "./figures"

splicing_antigenicity <- readRDS("./input_data/baseline_POST_gene_metric_mean_len_norm_no_gene_norm.rds")
top_genes <- unique(unname(unlist(lapply(seq(ncol(splicing_antigenicity)),function(col_val){
  genes <- rownames(splicing_antigenicity)
  SA_vec <- splicing_antigenicity[,col_val,drop=T]
  names(SA_vec)<-genes
  top_genes <- names(sort(SA_vec))[seq(nrow(splicing_antigenicity))]
  return(top_genes)
}))))

coding_potential <- readRDS("./input_data/baseline_POST_coding_potential_LGC.rds")
coding_potential_POS <- coding_potential %>% dplyr::filter(coding_potential > 0)
POS_CP_genes <- rownames(coding_potential_POS)
colnames_splicing_antigenicity <- colnames(splicing_antigenicity)
splicing_antigenicity_filt <- splicing_antigenicity[intersect(POS_CP_genes,top_genes),]


median_vals <- apply(splicing_antigenicity_filt,2,median,na.rm=T)
mean_vals <- apply(splicing_antigenicity_filt,2,mean,na.rm=T)
max_vals <- apply(splicing_antigenicity_filt,2,max,na.rm=T)
splicing_antigenicity_summ <- data.frame(median=median_vals,mean=mean_vals,max=max_vals)

manifest$SA_median <- vapply(manifest$Sample,function(samp){
  return(splicing_antigenicity_summ[samp,"median"])
},numeric(1))
manifest$SA_mean <- vapply(manifest$Sample,function(samp){
  return(splicing_antigenicity_summ[samp,"mean"])
},numeric(1))
manifest$SA_max <- vapply(manifest$Sample,function(samp){
  return(splicing_antigenicity_summ[samp,"max"])
},numeric(1))
manifest$tumor_purity <- vapply(manifest$Sample,function(samp){
  return(ESTIMATE_scores[samp,"tumor_purity"])
},numeric(1))
manifest$pseudo_purity <- vapply(manifest$Sample,function(samp){
  return(ESTIMATE_scores[samp,"pseudo_purity"])
},numeric(1))
manifest$ImmuneScore <- vapply(manifest$Sample,function(samp){
  return(as.numeric(ESTIMATE_scores[samp,"ImmuneScore"]))
},numeric(1))
manifest$StromalScore <- vapply(manifest$Sample,function(samp){
  return(as.numeric(ESTIMATE_scores[samp,"StromalScore"]))
},numeric(1))

manifest_no_factor <- manifest
manifest$AX_BOR3 <- factor(manifest$AX_BOR3,levels=c("CRPR","SD","PD"))

manifest_PRE <- manifest %>% dplyr::filter(AX_TIMETEMP=="PRE" & AX_BOR3!="NE")
manifest_PRE$TRTGRP <- vapply(manifest_PRE$TRTGRP,function(treat){
  if (str_detect(treat,"IPI")){
    return("NIV-IPI")
  } else {
    return(treat)
  }
},character(1))

png(sprintf("%s/%s",figure_dir,"FigS3B.png"),width = 480, height = 300)

manifest_PRE$SA_norm <- manifest_PRE$SA_mean/manifest_PRE$pseudo_purity
manifest_PRE_filt <- manifest_PRE %>% dplyr::filter(!is.infinite(SA_norm))
wilcox_test_val <- compare_means(SA_norm ~ AX_BOR3, p.adjust.method = "BH", data=manifest_PRE_filt,group.by = "TRTGRP")
cohens_d <- calculate_cohens_d(wilcox_test_val,c("TRTGRP"),"AX_BOR3",manifest_PRE_filt)
max_PRE_norm <- max(manifest_PRE_filt$SA_norm,na.rm=T)
wilcox_test_val$y.position <- rep(rep(max_PRE_norm,3)+seq(from=0,to=6,by=3),3)
wilcox_test_val$cohens_d <- format(round(cohens_d$cohens_d, 2), nsmall = 2)
wilcox_test_val <- wilcox_test_val %>% dplyr::filter(p.adj<0.05)
if (nrow(wilcox_test_val)==0){
  ggplot(manifest_PRE_filt,aes(x=AX_BOR3,y=SA_norm))+
  geom_boxplot(aes(fill=AX_BOR3))+facet_grid(.~TRTGRP)+geom_point()+
  ylab("SA norm. PRE-treatment")+
  xlab("Response")+
  theme(legend.position = "bottom",
        text = element_text(size = 20),
        axis.text.x=element_blank(),
        axis.title.x=element_blank())+
  guides(fill=guide_legend(title="Response"))+
  ylim(c(0,max_PRE_norm+14))

} else {
  ggplot(manifest_PRE_filt,aes(x=AX_BOR3,y=SA_norm))+
    geom_boxplot(aes(fill=AX_BOR3))+facet_grid(.~TRTGRP)+geom_point()+
    add_pvalue(wilcox_test_val,label="d={cohens_d}\n{p.signif}", remove.bracket = FALSE,label.size=6) +
    ylab("SA norm. PRE-treatment")+
    xlab("Response")+
    theme(legend.position = "bottom",
          text = element_text(size = 20),
          axis.text.x=element_blank(),
          axis.title.x=element_blank())+
    guides(fill=guide_legend(title="Response"))+
  ylim(c(0,max_PRE_norm+14))

}

dev.off()

png(sprintf("%s/%s",figure_dir,"FigS3C.png"),width = 480/1.5, height = 300/1.25)

wilcox_test_val <- compare_means(SA_norm ~ AX_BOR3, p.adjust.method = "BH", data=manifest_PRE_filt)
max_PRE_norm <- max(manifest_PRE_filt$SA_norm,na.rm=T)
wilcox_test_val$y.position <- rep(max_PRE_norm,3)+seq(from=3,to=9,by=3)
cohens_d <- calculate_cohens_d(wilcox_test_val,c(),"AX_BOR3",manifest_PRE_filt)
wilcox_test_val$cohens_d <- format(round(cohens_d$cohens_d, 2), nsmall = 2)
wilcox_test_val <- wilcox_test_val %>% dplyr::filter(p<0.05)
if (nrow(wilcox_test_val)==0){
  ggplot(manifest_PRE_filt,aes(x=AX_BOR3,y=SA_norm))+
    geom_boxplot(aes(fill=AX_BOR3))+geom_point()+
    add_pvalue(wilcox_test_val,label="d={cohens_d}\n{p.signif}", remove.bracket = FALSE,label.size=6)+
    ylab("SA normalized PRE treatment")+
    xlab("Response")+
    theme(legend.position = "right",
          text = element_text(size = 20),
          axis.text.x=element_blank(),
          axis.title.x=element_blank())+
    guides(fill=guide_legend(title="Response"))+
  ylim(c(0,max_PRE_norm+14))
  
} else {
  ggplot(manifest_PRE_filt,aes(x=AX_BOR3,y=SA_norm))+
    geom_boxplot(aes(fill=AX_BOR3))+geom_point()+
    add_pvalue(wilcox_test_val, label = "d={cohens_d}\n{p.signif}", remove.bracket = FALSE,label.size=6)+
    ylab("SA norm. PRE treatment")+
    xlab("Response")+
    theme(legend.position = "right",
          text = element_text(size = 15),
          axis.text.x=element_blank(),
          axis.title.x=element_blank())+
    guides(fill=guide_legend(title="Response"))+
  ylim(c(0,max_PRE_norm+14))
  
}

dev.off()

# png(sprintf("%s/%s",figure_dir,"SA_mean_PRE.png"),width = 480, height = 300)
# 
# manifest_PRE_filt <- manifest_PRE %>% dplyr::filter(!is.infinite(SA_norm))
# wilcox_test_val <- compare_means(SA_mean ~ AX_BOR3, p.adjust.method = "BH", data=manifest_PRE_filt,group.by = "TRTGRP")
# max_PRE_mean <- max(manifest_PRE_filt$SA_mean,na.rm=T)
# wilcox_test_val$y.position <- rep(rep(max_PRE_mean,3)+seq(from=1,to=3,by=1),3)
# cohens_d <- calculate_cohens_d(wilcox_test_val,c("TRTGRP"),"AX_BOR3",manifest_PRE_filt)
# wilcox_test_val$cohens_d <- format(round(cohens_d$cohens_d, 2), nsmall = 2)
# wilcox_test_val <- wilcox_test_val %>% dplyr::filter(p.adj<0.05)
# 
# if (nrow(wilcox_test_val)==0){
#   ggplot(manifest_PRE_filt,aes(x=AX_BOR3,y=SA_mean))+
#     geom_boxplot(aes(fill=AX_BOR3))+facet_grid(.~TRTGRP)+geom_point()+
#     ylab("SA mean PRE-treatment")+
#     xlab("Response")+
#     theme(legend.position = "bottom",
#           text = element_text(size = 20),
#           axis.text.x=element_blank(),
#           axis.title.x=element_blank())+
#     guides(fill=guide_legend(title="Response"))
# } else {
#   ggplot(manifest_PRE_filt,aes(x=AX_BOR3,y=SA_mean))+
#     geom_boxplot(aes(fill=AX_BOR3))+facet_grid(.~TRTGRP)+geom_point()+
#     add_pvalue(wilcox_test_val, label = "p={p.adj},d={cohens_d}", remove.bracket = FALSE,label.size=4)+
#     ylab("SA mean PRE-treatment")+
#     xlab("Response")+
#     theme(legend.position = "bottom",
#           text = element_text(size = 20),
#           axis.text.x=element_blank(),
#           axis.title.x=element_blank())+
#     guides(fill=guide_legend(title="Response"))
# }
# 
# dev.off()

# png(sprintf("%s/%s",figure_dir,"SA_mean_PRE_all.png"),width = 480/1.5, height = 300/1.25)
# 
# wilcox_test_val <- compare_means(SA_mean ~ AX_BOR3, p.adjust.method = "BH", data=manifest_PRE_filt)
# wilcox_test_val$y.position <- rep(max_PRE_mean,3)+seq(from=1,to=3,by=1)
# cohens_d <- calculate_cohens_d(wilcox_test_val,c(),"AX_BOR3",manifest_PRE_filt)
# wilcox_test_val$cohens_d <- format(round(cohens_d$cohens_d, 2), nsmall = 2)
# wilcox_test_val <- wilcox_test_val %>% dplyr::filter(p<0.05)
# if (nrow(wilcox_test_val)==0){
#   ggplot(manifest_PRE_filt,aes(x=AX_BOR3,y=SA_mean))+
#     geom_boxplot(aes(fill=AX_BOR3))+geom_point()+
#     ylab("SA mean PRE-treatment")+
#     xlab("Response")+
#     theme(legend.position = "right",
#           text = element_text(size = 20),
#           axis.text.x=element_blank(),
#           axis.title.x=element_blank())+
#     guides(fill=guide_legend(title="Response"))
# } else {
#   ggplot(manifest_PRE_filt,aes(x=AX_BOR3,y=SA_mean))+
#     geom_boxplot(aes(fill=AX_BOR3))+geom_point()+
#     ylab("SA mean PRE-treatment")+
#     xlab("Response")+
#     theme(legend.position = "right",
#           text = element_text(size = 20),
#           axis.text.x=element_blank(),
#           axis.title.x=element_blank())+
#     guides(fill=guide_legend(title="Response"))
# }
# 
# dev.off()

# png(sprintf("%s/%s",figure_dir,"PP_PRE.png"),width = 480, height = 300)
# 
# wilcox_test_val <- compare_means(pseudo_purity ~ AX_BOR3, p.adjust.method = "BH", data=manifest_PRE_filt,group.by = "TRTGRP")
# max_pseudo_purity <- max(manifest_PRE_filt$pseudo_purity,na.rm=T)
# wilcox_test_val$y.position <- rep(rep(max_pseudo_purity,3)+seq(from=max_pseudo_purity/6,to=(max_pseudo_purity/6)*3,by=max_pseudo_purity/6),3)
# cohens_d <- calculate_cohens_d(wilcox_test_val,c("TRTGRP"),"AX_BOR3",manifest_PRE_filt)
# wilcox_test_val$cohens_d <- format(round(cohens_d$cohens_d, 2), nsmall = 2)
# wilcox_test_val <- wilcox_test_val %>% dplyr::filter(p.adj<0.05)
# if (nrow(wilcox_test_val)==0){
#   ggplot(manifest_PRE_filt,aes(x=AX_BOR3,y=pseudo_purity))+
#     geom_boxplot(aes(fill=AX_BOR3))+geom_point()+facet_grid(.~TRTGRP)+
#     ylab("Pseudo Purity PRE-treatment")+
#     xlab("Response")+
#     theme(legend.position = "bottom",
#           text = element_text(size = 20),
#           axis.text.x=element_blank(),
#           axis.title.x=element_blank())+
#     guides(fill=guide_legend(title="Response"))
# } else {
#     ggplot(manifest_PRE_filt,aes(x=AX_BOR3,y=pseudo_purity))+
#     geom_boxplot(aes(fill=AX_BOR3))+geom_point()+facet_grid(.~TRTGRP)+
#     add_pvalue(wilcox_test_val, label = "p={p.adj},d={cohens_d}", remove.bracket = FALSE,label.size=4)+
#     ylab("Pseudo Purity PRE-treatment")+
#     xlab("Response")+
#     theme(legend.position = "bottom",
#           text = element_text(size = 20),
#           axis.text.x=element_blank(),
#           axis.title.x=element_blank())+
#     guides(fill=guide_legend(title="Response"))
# }
# 
# dev.off()

# png(sprintf("%s/%s",figure_dir,"PP_PRE_all.png"),width = 480/1.5, height = 300/1.25)
# 
# wilcox_test_val <- compare_means(pseudo_purity ~ AX_BOR3, p.adjust.method = "BH", data=manifest_PRE_filt)
# max_pseudo_purity <- max(manifest_PRE_filt$pseudo_purity,na.rm=T)
# wilcox_test_val$y.position <- rep(max_pseudo_purity,3)+seq(from=max_pseudo_purity/6,to=(max_pseudo_purity/6)*3,by=max_pseudo_purity/6)
# cohens_d <- calculate_cohens_d(wilcox_test_val,c(),"AX_BOR3",manifest_PRE_filt)
# wilcox_test_val$cohens_d <- format(round(cohens_d$cohens_d, 2), nsmall = 2)
# wilcox_test_val <- wilcox_test_val %>% dplyr::filter(p<0.05)
# if (nrow(wilcox_test_val)==0){
#   ggplot(manifest_PRE_filt,aes(x=AX_BOR3,y=pseudo_purity))+
#     geom_boxplot(aes(fill=AX_BOR3))+geom_point()+
#     ylab("Pseudo Purity PRE-treatment")+
#     xlab("Response")+
#     theme(legend.position = "right",
#           text = element_text(size = 20),
#           axis.text.x=element_blank(),
#           axis.title.x=element_blank())+
#     guides(fill=guide_legend(title="Response"))
# } else {
#   ggplot(manifest_PRE_filt,aes(x=AX_BOR3,y=pseudo_purity))+
#     geom_boxplot(aes(fill=AX_BOR3))+geom_point()+
#     add_pvalue(wilcox_test_val, label = "p={p.format},d={cohens_d}", remove.bracket = FALSE,label.size=4)+
#     ylab("Pseudo Purity PRE-treatment")+
#     xlab("Response")+
#     theme(legend.position = "right",
#           text = element_text(size = 20),
#           axis.text.x=element_blank(),
#           axis.title.x=element_blank())+
#     guides(fill=guide_legend(title="Response"))
# }
# 
# dev.off()

# png(sprintf("%s/%s",figure_dir,"IM_PRE.png"),width = 480, height = 300)
# 
# wilcox_test_val <- compare_means(ImmuneScore ~ AX_BOR3, p.adjust.method = "BH", data=manifest_PRE_filt,group.by = "TRTGRP")
# max_imm <- max(manifest_PRE_filt$ImmuneScore,na.rm=T)
# wilcox_test_val$y.position <- rep(rep(max_imm,3)+seq(from=max_imm/6,to=(max_imm/6)*3,by=max_imm/6),3)
# cohens_d <- calculate_cohens_d(wilcox_test_val,c("TRTGRP"),"AX_BOR3",manifest_PRE_filt)
# wilcox_test_val$cohens_d <- format(round(cohens_d$cohens_d, 2), nsmall = 2)
# wilcox_test_val <- wilcox_test_val %>% dplyr::filter(p.adj<0.05)
# if (nrow(wilcox_test_val)==0){
#   ggplot(manifest_PRE_filt,aes(x=AX_BOR3,y=ImmuneScore))+
#     geom_boxplot(aes(fill=AX_BOR3))+geom_point()+facet_grid(.~TRTGRP)+
#     ylab("Imm. Score PRE-treatment")+
#     xlab("Response")+
#     theme(legend.position = "bottom",
#           text = element_text(size = 20),
#           axis.text.x=element_blank(),
#           axis.title.x=element_blank())+
#     guides(fill=guide_legend(title="Response"))
# } else {
#   ggplot(manifest_PRE_filt,aes(x=AX_BOR3,y=ImmuneScore))+
#     geom_boxplot(aes(fill=AX_BOR3))+geom_point()+facet_grid(.~TRTGRP)+
#     add_pvalue(wilcox_test_val, label = "p={p.adj},d={cohens_d}", remove.bracket = FALSE,label.size=4)+
#     ylab("Imm. Score PRE-treatment")+
#     xlab("Response")+
#     theme(legend.position = "bottom",
#           text = element_text(size = 20),
#           axis.text.x=element_blank(),
#           axis.title.x=element_blank())+
#     guides(fill=guide_legend(title="Response"))
# }
# 
# dev.off()

# png(sprintf("%s/%s",figure_dir,"IM_PRE_all.png"),width = 480/1.5, height = 300/1.25)
# 
# wilcox_test_val <- compare_means(ImmuneScore ~ AX_BOR3, p.adjust.method = "BH", data=manifest_PRE_filt)
# max_imm <- max(manifest_PRE_filt$ImmuneScore,na.rm=T)
# wilcox_test_val$y.position <- rep(max_imm,3)+seq(from=max_imm/6,to=(max_imm/6)*3,by=max_imm/6)
# cohens_d <- calculate_cohens_d(wilcox_test_val,c(),"AX_BOR3",manifest_PRE_filt)
# wilcox_test_val$cohens_d <- format(round(cohens_d$cohens_d, 2), nsmall = 2)
# wilcox_test_val <- wilcox_test_val %>% dplyr::filter(p<0.05)
# if (nrow(wilcox_test_val)==0){
# ggplot(manifest_PRE_filt,aes(x=AX_BOR3,y=ImmuneScore))+
#   geom_boxplot(aes(fill=AX_BOR3))+geom_point()+
#   ylab("Imm. Score PRE-treatment")+
#   xlab("Response")+
#   theme(legend.position = "right",
#         text = element_text(size = 20),
#         axis.text.x=element_blank(),
#         axis.title.x=element_blank())+
#   guides(fill=guide_legend(title="Response"))
# } else {
#   ggplot(manifest_PRE_filt,aes(x=AX_BOR3,y=ImmuneScore))+
#     geom_boxplot(aes(fill=AX_BOR3))+geom_point()+
#     add_pvalue(wilcox_test_val, label = "p={p.format},d={cohens_d}", remove.bracket = FALSE,label.size=4)+
#     ylab("Imm. Score PRE-treatment")+
#     xlab("Response")+
#     theme(legend.position = "right",
#           text = element_text(size = 20),
#           axis.text.x=element_blank(),
#           axis.title.x=element_blank())+
#     guides(fill=guide_legend(title="Response"))
# }
# 
# dev.off()

# png(sprintf("%s/%s",figure_dir,"ST_PRE.png"),width = 480, height = 300)
# 
# wilcox_test_val <- compare_means(StromalScore ~ AX_BOR3, p.adjust.method = "BH", data=manifest_PRE_filt,group.by = "TRTGRP")
# max_strom <- max(manifest_PRE_filt$StromalScore,na.rm=T)
# wilcox_test_val$y.position <- rep(rep(max_strom,3)+seq(from=max_strom/6,to=(max_strom/6)*3,by=max_strom/6),3)
# cohens_d <- calculate_cohens_d(wilcox_test_val,c("TRTGRP"),"AX_BOR3",manifest_PRE_filt)
# wilcox_test_val$cohens_d <- format(round(cohens_d$cohens_d, 2), nsmall = 2)
# wilcox_test_val <- wilcox_test_val %>% dplyr::filter(p.adj<0.05)
# if (nrow(wilcox_test_val)==0){
#   ggplot(manifest_PRE_filt,aes(x=AX_BOR3,y=StromalScore))+
#     geom_boxplot(aes(fill=AX_BOR3))+geom_point()+facet_grid(.~TRTGRP)+
#     ylab("Strom. Score PRE-treatment")+
#     xlab("Response")+
#     theme(legend.position = "bottom",
#           text = element_text(size = 20),
#           axis.text.x=element_blank(),
#           axis.title.x=element_blank())+
#     guides(fill=guide_legend(title="Response"))
# } else {
#   ggplot(manifest_PRE_filt,aes(x=AX_BOR3,y=StromalScore))+
#     geom_boxplot(aes(fill=AX_BOR3))+geom_point()+facet_grid(.~TRTGRP)+
#     add_pvalue(wilcox_test_val, label = "p={p.adj},d={cohens_d}", remove.bracket = FALSE,label.size=4)+
#     ylab("Strom. Score PRE-treatment")+
#     xlab("Response")+
#     theme(legend.position = "bottom",
#           text = element_text(size = 20),
#           axis.text.x=element_blank(),
#           axis.title.x=element_blank())+
#     guides(fill=guide_legend(title="Response"))
# }
# 
# dev.off()

# png(sprintf("%s/%s",figure_dir,"ST_PRE_all.png"),width = 480/1.5, height = 300/1.25)
# 
# wilcox_test_val <- compare_means(StromalScore ~ AX_BOR3, p.adjust.method = "BH", data=manifest_PRE_filt)
# max_strom <- max(manifest_PRE_filt$StromalScore,na.rm=T)
# wilcox_test_val$y.position <- rep(max_strom,3)+seq(from=max_strom/6,to=(max_strom/6)*3,by=max_strom/6)
# cohens_d <- calculate_cohens_d(wilcox_test_val,c(),"AX_BOR3",manifest_PRE_filt)
# wilcox_test_val$cohens_d <- format(round(cohens_d$cohens_d, 2), nsmall = 2)
# wilcox_test_val <- wilcox_test_val %>% dplyr::filter(p<0.05)
# if (nrow(wilcox_test_val)==0){
# ggplot(manifest_PRE_filt,aes(x=AX_BOR3,y=StromalScore))+
#   geom_boxplot(aes(fill=AX_BOR3))+geom_point()+
#   ylab("Strom. Score PRE-treatment")+
#   xlab("Response")+
#   theme(legend.position = "right",
#         text = element_text(size = 20),
#         axis.text.x=element_blank(),
#         axis.title.x=element_blank())+
#   guides(fill=guide_legend(title="Response"))
# 
# } else {
#   ggplot(manifest_PRE_filt,aes(x=AX_BOR3,y=StromalScore))+
#     geom_boxplot(aes(fill=AX_BOR3))+geom_point()+
#     add_pvalue(wilcox_test_val, label = "p={p.format},d={cohens_d}", remove.bracket = FALSE,label.size=4)+
#     ylab("Strom. Score PRE-treatment")+
#     xlab("Response")+
#     theme(legend.position = "right",
#           text = element_text(size = 20),
#           axis.text.x=element_blank(),
#           axis.title.x=element_blank())+
#     guides(fill=guide_legend(title="Response"))
# }
# 
# dev.off()

# png(sprintf("%s/%s",figure_dir,"SA_mean_PP_PRE_all.png"),width = 480, height = 300)
# 
# ggplot(manifest_PRE,aes(x=pseudo_purity,y=SA_mean))+geom_point()+
#   stat_cor(method = "kendall",label.x.npc="middle",label.y.npc="top",size = 5)+
#   geom_smooth(method="lm",se=F)+
#   theme(text = element_text(size = 20))
# 
# dev.off()

# png(sprintf("%s/%s",figure_dir,"SA_mean_PP_PRE.png"),width = 480, height = 300)
# 
# ggplot(manifest_PRE,aes(x=pseudo_purity,y=SA_mean))+geom_point()+
#   stat_cor(method = "kendall",label.x.npc="left",label.y.npc="top",size = 4)+
#   geom_smooth(method="lm",se=F)+facet_grid(AX_BOR3~TRTGRP)+
#   theme(text = element_text(size = 20))
# 
# dev.off()

png(sprintf("%s/%s",figure_dir,"FigS3A.png"),width = 480, height = 300)

ggplot(manifest_PRE,aes(x=pseudo_purity,y=SA_mean))+geom_point()+
  stat_cor(method = "kendall",label.x.npc="left",label.y.npc="top",size = 5)+
  geom_smooth(method="lm",se=F)+facet_grid(AX_BOR3~.)+
  theme(text = element_text(size = 20))

dev.off()

# png(sprintf("%s/%s",figure_dir,"SA_mean_IM_PRE_all.png"),width = 480, height = 300)
# 
# ggplot(manifest_PRE,aes(x=ImmuneScore,y=SA_mean))+geom_point()+
#   stat_cor(method = "kendall",label.x.npc="middle",label.y.npc="top",size = 5)+
#   geom_smooth(method="lm",se=F)+
#   theme(text = element_text(size = 20))
# 
# dev.off()

# png(sprintf("%s/%s",figure_dir,"SA_mean_IM_PRE.png"),width = 480, height = 300)
# 
# ggplot(manifest_PRE,aes(x=ImmuneScore,y=SA_mean))+geom_point()+
#   stat_cor(method = "kendall",label.x.npc="left",label.y.npc="top",size = 4)+
#   geom_smooth(method="lm",se=F)+facet_grid(AX_BOR3~TRTGRP)+
#   theme(text = element_text(size = 20))
# 
# dev.off()

# png(sprintf("%s/%s",figure_dir,"SA_mean_IM_PRE_resp.png"),width = 480, height = 300)
# 
# ggplot(manifest_PRE,aes(x=ImmuneScore,y=SA_mean))+geom_point()+
#   stat_cor(method = "kendall",label.x.npc="left",label.y.npc="top",size = 5)+
#   geom_smooth(method="lm",se=F)+facet_grid(AX_BOR3~.)+
#   theme(text = element_text(size = 20))
# 
# dev.off()

# png(sprintf("%s/%s",figure_dir,"SA_mean_ST_PRE_all.png"),width = 480, height = 300)
# 
# ggplot(manifest_PRE,aes(x=StromalScore,y=SA_mean))+geom_point()+
#   stat_cor(method = "kendall",label.x.npc="middle",label.y.npc="top",size = 5)+
#   geom_smooth(method="lm",se=F)+
#   theme(text = element_text(size = 20))
# 
# dev.off()

# png(sprintf("%s/%s",figure_dir,"SA_mean_ST_PRE.png"),width = 480, height = 300)
# 
# ggplot(manifest_PRE,aes(x=StromalScore,y=SA_mean))+geom_point()+
#   stat_cor(method = "kendall",label.x.npc="left",label.y.npc="top",size = 5)+
#   geom_smooth(method="lm",se=F)+facet_grid(AX_BOR3~TRTGRP)+
#   theme(text = element_text(size = 20))
# 
# dev.off()

# png(sprintf("%s/%s",figure_dir,"SA_mean_ST_PRE_resp.png"),width = 480, height = 300)
# 
# ggplot(manifest_PRE,aes(x=StromalScore,y=SA_mean))+geom_point()+
#   stat_cor(method = "kendall",label.x.npc="left",label.y.npc="top",size = 5)+
#   geom_smooth(method="lm",se=F)+facet_grid(AX_BOR3~.)+
#   theme(text = element_text(size = 20))
# 
# dev.off()

# png(sprintf("%s/%s",figure_dir,"PP_TP_PRE_all.png"),width = 480, height = 300)
# 
# ggplot(manifest_PRE,aes(x=tumor_purity,y=pseudo_purity))+geom_point()+
#   stat_cor(method = "kendall",label.x.npc="middle",label.y.npc="top",size = 5)+
#   geom_smooth(method="lm",se=F)+
#   theme(text = element_text(size = 20))
# 
# dev.off()

```

# Evaluating the splicing antigenicity POST treatment

```{r}

figure_dir <- "./figures"

splicing_antigenicity <- readRDS("./input_data/baseline_POST_gene_metric_mean_len_norm_no_gene_norm.rds")
top_genes <- unique(unname(unlist(lapply(seq(ncol(splicing_antigenicity)),function(col_val){
  genes <- rownames(splicing_antigenicity)
  SA_vec <- splicing_antigenicity[,col_val,drop=T]
  names(SA_vec)<-genes
  top_genes <- names(sort(SA_vec))[seq(nrow(splicing_antigenicity))]
  return(top_genes)
}))))

coding_potential <- readRDS("./input_data/baseline_POST_coding_potential_LGC.rds")
coding_potential_POS <- coding_potential %>% dplyr::filter(coding_potential > 0)
POS_CP_genes <- rownames(coding_potential_POS)
colnames_splicing_antigenicity <- colnames(splicing_antigenicity)
splicing_antigenicity_filt <- splicing_antigenicity[intersect(POS_CP_genes,top_genes),]


median_vals <- apply(splicing_antigenicity_filt,2,median,na.rm=T)
mean_vals <- apply(splicing_antigenicity_filt,2,mean,na.rm=T)
max_vals <- apply(splicing_antigenicity_filt,2,max,na.rm=T)
splicing_antigenicity_summ <- data.frame(median=median_vals,mean=mean_vals,max=max_vals)

manifest$SA_median <- vapply(manifest$Sample,function(samp){
  return(splicing_antigenicity_summ[samp,"median"])
},numeric(1))
manifest$SA_mean <- vapply(manifest$Sample,function(samp){
  return(splicing_antigenicity_summ[samp,"mean"])
},numeric(1))
manifest$SA_max <- vapply(manifest$Sample,function(samp){
  return(splicing_antigenicity_summ[samp,"max"])
},numeric(1))
manifest$tumor_purity <- vapply(manifest$Sample,function(samp){
  return(ESTIMATE_scores[samp,"tumor_purity"])
},numeric(1))
manifest$pseudo_purity <- vapply(manifest$Sample,function(samp){
  return(ESTIMATE_scores[samp,"pseudo_purity"])
},numeric(1))
manifest$ImmuneScore <- vapply(manifest$Sample,function(samp){
  return(as.numeric(ESTIMATE_scores[samp,"ImmuneScore"]))
},numeric(1))
manifest$StromalScore <- vapply(manifest$Sample,function(samp){
  return(as.numeric(ESTIMATE_scores[samp,"StromalScore"]))
},numeric(1))

manifest_no_factor <- manifest
manifest$AX_BOR3 <- factor(manifest$AX_BOR3,levels=c("CRPR","SD","PD"))

manifest_POST <- manifest %>% dplyr::filter(AX_TIMETEMP=="POST" & AX_BOR3!="NE")
manifest_POST$TRTGRP <- vapply(manifest_POST$TRTGRP,function(treat){
  if (str_detect(treat,"IPI")){
    return("NIV-IPI")
  } else {
    return(treat)
  }
},character(1))

png(sprintf("%s/%s",figure_dir,"Fig5B.png"),width = 480, height = 300)

manifest_POST$SA_norm <- manifest_POST$SA_mean/manifest_POST$pseudo_purity
manifest_POST_filt <- manifest_POST %>% dplyr::filter(!is.infinite(SA_norm))
wilcox_test_val <- compare_means(SA_norm ~ AX_BOR3, p.adjust.method = "BH", data=manifest_POST_filt,group.by = "TRTGRP")
cohens_d <- calculate_cohens_d(wilcox_test_val,c("TRTGRP"),"AX_BOR3",manifest_POST_filt)
max_POST_norm <- max(manifest_POST_filt$SA_norm,na.rm=T)
wilcox_test_val$y.position <- rep(rep(max_POST_norm,3)+seq(from=0,to=6,by=3),3)
wilcox_test_val$cohens_d <- format(round(cohens_d$cohens_d, 2), nsmall = 2)
wilcox_test_val <- wilcox_test_val %>% dplyr::filter(p.adj<0.05)
if (nrow(wilcox_test_val)==0){
  ggplot(manifest_POST_filt,aes(x=AX_BOR3,y=SA_norm))+
  geom_boxplot(aes(fill=AX_BOR3))+facet_grid(.~TRTGRP)+geom_point()+
  ylab("SA norm. POST-treatment")+
  xlab("Response")+
  theme(legend.position = "bottom",
        text = element_text(size = 20),
        axis.text.x=element_blank(),
        axis.title.x=element_blank())+
  guides(fill=guide_legend(title="Response"))+
  ylim(c(0,max_POST_norm+14))

} else {
  ggplot(manifest_POST_filt,aes(x=AX_BOR3,y=SA_norm))+
    geom_boxplot(aes(fill=AX_BOR3))+facet_grid(.~TRTGRP)+geom_point()+
    add_pvalue(wilcox_test_val,label="d={cohens_d}\n{p.signif}", remove.bracket = FALSE,label.size=6) +
    ylab("SA norm. POST-treatment")+
    xlab("Response")+
    theme(legend.position = "bottom",
          text = element_text(size = 20),
          axis.text.x=element_blank(),
          axis.title.x=element_blank())+
    guides(fill=guide_legend(title="Response"))+
  ylim(c(0,max_POST_norm+14))

}

dev.off()

png(sprintf("%s/%s",figure_dir,"Fig5C.png"),width = 480/1.5, height = 300/1.25)

wilcox_test_val <- compare_means(SA_norm ~ AX_BOR3, p.adjust.method = "BH", data=manifest_POST_filt)
max_POST_norm <- max(manifest_POST_filt$SA_norm,na.rm=T)
wilcox_test_val$y.position <- rep(max_POST_norm,3)+seq(from=3,to=9,by=3)
cohens_d <- calculate_cohens_d(wilcox_test_val,c(),"AX_BOR3",manifest_POST_filt)
wilcox_test_val$cohens_d <- format(round(cohens_d$cohens_d, 2), nsmall = 2)
wilcox_test_val <- wilcox_test_val %>% dplyr::filter(p<0.05)
if (nrow(wilcox_test_val)==0){
  ggplot(manifest_POST_filt,aes(x=AX_BOR3,y=SA_norm))+
    geom_boxplot(aes(fill=AX_BOR3))+geom_point()+
    add_pvalue(wilcox_test_val,label="d={cohens_d}\n{p.signif}", remove.bracket = FALSE,label.size=6)+
    ylab("SA normalized POST treatment")+
    xlab("Response")+
    theme(legend.position = "right",
          text = element_text(size = 20),
          axis.text.x=element_blank(),
          axis.title.x=element_blank())+
    guides(fill=guide_legend(title="Response"))+
  ylim(c(0,max_POST_norm+14))
  
} else {
  ggplot(manifest_POST_filt,aes(x=AX_BOR3,y=SA_norm))+
    geom_boxplot(aes(fill=AX_BOR3))+geom_point()+
    add_pvalue(wilcox_test_val, label = "d={cohens_d}\n{p.signif}", remove.bracket = FALSE,label.size=6)+
    ylab("SA norm. POST treatment")+
    xlab("Response")+
    theme(legend.position = "right",
          text = element_text(size = 15),
          axis.text.x=element_blank(),
          axis.title.x=element_blank())+
    guides(fill=guide_legend(title="Response"))+
  ylim(c(0,max_POST_norm+14))
  
}

dev.off()

# png(sprintf("%s/%s",figure_dir,"SA_mean_POST.png"),width = 480, height = 300)
# 
# manifest_POST_filt <- manifest_POST %>% dplyr::filter(!is.infinite(SA_norm))
# wilcox_test_val <- compare_means(SA_mean ~ AX_BOR3, p.adjust.method = "BH", data=manifest_POST_filt,group.by = "TRTGRP")
# max_POST_mean <- max(manifest_POST_filt$SA_mean,na.rm=T)
# wilcox_test_val$y.position <- rep(rep(max_POST_mean,3)+seq(from=1,to=3,by=1),3)
# cohens_d <- calculate_cohens_d(wilcox_test_val,c("TRTGRP"),"AX_BOR3",manifest_POST_filt)
# wilcox_test_val$cohens_d <- format(round(cohens_d$cohens_d, 2), nsmall = 2)
# wilcox_test_val <- wilcox_test_val %>% dplyr::filter(p.adj<0.05)
# 
# if (nrow(wilcox_test_val)==0){
#   ggplot(manifest_POST_filt,aes(x=AX_BOR3,y=SA_mean))+
#     geom_boxplot(aes(fill=AX_BOR3))+facet_grid(.~TRTGRP)+geom_point()+
#     ylab("SA mean POST-treatment")+
#     xlab("Response")+
#     theme(legend.position = "bottom",
#           text = element_text(size = 20),
#           axis.text.x=element_blank(),
#           axis.title.x=element_blank())+
#     guides(fill=guide_legend(title="Response"))
# } else {
#   ggplot(manifest_POST_filt,aes(x=AX_BOR3,y=SA_mean))+
#     geom_boxplot(aes(fill=AX_BOR3))+facet_grid(.~TRTGRP)+geom_point()+
#     add_pvalue(wilcox_test_val, label = "p={p.adj},d={cohens_d}", remove.bracket = FALSE,label.size=4)+
#     ylab("SA mean POST-treatment")+
#     xlab("Response")+
#     theme(legend.position = "bottom",
#           text = element_text(size = 20),
#           axis.text.x=element_blank(),
#           axis.title.x=element_blank())+
#     guides(fill=guide_legend(title="Response"))
# }
# 
# dev.off()

# png(sprintf("%s/%s",figure_dir,"SA_mean_POST_all.png"),width = 480/1.5, height = 300/1.25)
# 
# wilcox_test_val <- compare_means(SA_mean ~ AX_BOR3, p.adjust.method = "BH", data=manifest_POST_filt)
# wilcox_test_val$y.position <- rep(max_POST_mean,3)+seq(from=1,to=3,by=1)
# cohens_d <- calculate_cohens_d(wilcox_test_val,c(),"AX_BOR3",manifest_POST_filt)
# wilcox_test_val$cohens_d <- format(round(cohens_d$cohens_d, 2), nsmall = 2)
# wilcox_test_val <- wilcox_test_val %>% dplyr::filter(p<0.05)
# if (nrow(wilcox_test_val)==0){
#   ggplot(manifest_POST_filt,aes(x=AX_BOR3,y=SA_mean))+
#     geom_boxplot(aes(fill=AX_BOR3))+geom_point()+
#     ylab("SA mean POST-treatment")+
#     xlab("Response")+
#     theme(legend.position = "right",
#           text = element_text(size = 20),
#           axis.text.x=element_blank(),
#           axis.title.x=element_blank())+
#     guides(fill=guide_legend(title="Response"))
# } else {
#   ggplot(manifest_POST_filt,aes(x=AX_BOR3,y=SA_mean))+
#     geom_boxplot(aes(fill=AX_BOR3))+geom_point()+
#     ylab("SA mean POST-treatment")+
#     xlab("Response")+
#     theme(legend.position = "right",
#           text = element_text(size = 20),
#           axis.text.x=element_blank(),
#           axis.title.x=element_blank())+
#     guides(fill=guide_legend(title="Response"))
# }
# 
# dev.off()

# png(sprintf("%s/%s",figure_dir,"PP_POST.png"),width = 480, height = 300)
# 
# wilcox_test_val <- compare_means(pseudo_purity ~ AX_BOR3, p.adjust.method = "BH", data=manifest_POST_filt,group.by = "TRTGRP")
# max_pseudo_purity <- max(manifest_POST_filt$pseudo_purity,na.rm=T)
# wilcox_test_val$y.position <- rep(rep(max_pseudo_purity,3)+seq(from=max_pseudo_purity/6,to=(max_pseudo_purity/6)*3,by=max_pseudo_purity/6),3)
# cohens_d <- calculate_cohens_d(wilcox_test_val,c("TRTGRP"),"AX_BOR3",manifest_POST_filt)
# wilcox_test_val$cohens_d <- format(round(cohens_d$cohens_d, 2), nsmall = 2)
# wilcox_test_val <- wilcox_test_val %>% dplyr::filter(p.adj<0.05)
# if (nrow(wilcox_test_val)==0){
#   ggplot(manifest_POST_filt,aes(x=AX_BOR3,y=pseudo_purity))+
#     geom_boxplot(aes(fill=AX_BOR3))+geom_point()+facet_grid(.~TRTGRP)+
#     ylab("Pseudo Purity POST-treatment")+
#     xlab("Response")+
#     theme(legend.position = "bottom",
#           text = element_text(size = 20),
#           axis.text.x=element_blank(),
#           axis.title.x=element_blank())+
#     guides(fill=guide_legend(title="Response"))
# } else {
#     ggplot(manifest_POST_filt,aes(x=AX_BOR3,y=pseudo_purity))+
#     geom_boxplot(aes(fill=AX_BOR3))+geom_point()+facet_grid(.~TRTGRP)+
#     add_pvalue(wilcox_test_val, label = "p={p.adj},d={cohens_d}", remove.bracket = FALSE,label.size=4)+
#     ylab("Pseudo Purity POST-treatment")+
#     xlab("Response")+
#     theme(legend.position = "bottom",
#           text = element_text(size = 20),
#           axis.text.x=element_blank(),
#           axis.title.x=element_blank())+
#     guides(fill=guide_legend(title="Response"))
# }
# 
# dev.off()

# png(sprintf("%s/%s",figure_dir,"PP_POST_all.png"),width = 480/1.5, height = 300/1.25)
# 
# wilcox_test_val <- compare_means(pseudo_purity ~ AX_BOR3, p.adjust.method = "BH", data=manifest_POST_filt)
# max_pseudo_purity <- max(manifest_POST_filt$pseudo_purity,na.rm=T)
# wilcox_test_val$y.position <- rep(max_pseudo_purity,3)+seq(from=max_pseudo_purity/6,to=(max_pseudo_purity/6)*3,by=max_pseudo_purity/6)
# cohens_d <- calculate_cohens_d(wilcox_test_val,c(),"AX_BOR3",manifest_POST_filt)
# wilcox_test_val$cohens_d <- format(round(cohens_d$cohens_d, 2), nsmall = 2)
# wilcox_test_val <- wilcox_test_val %>% dplyr::filter(p<0.05)
# if (nrow(wilcox_test_val)==0){
#   ggplot(manifest_POST_filt,aes(x=AX_BOR3,y=pseudo_purity))+
#     geom_boxplot(aes(fill=AX_BOR3))+geom_point()+
#     ylab("Pseudo Purity POST-treatment")+
#     xlab("Response")+
#     theme(legend.position = "right",
#           text = element_text(size = 20),
#           axis.text.x=element_blank(),
#           axis.title.x=element_blank())+
#     guides(fill=guide_legend(title="Response"))
# } else {
#   ggplot(manifest_POST_filt,aes(x=AX_BOR3,y=pseudo_purity))+
#     geom_boxplot(aes(fill=AX_BOR3))+geom_point()+
#     add_pvalue(wilcox_test_val, label = "p={p.format},d={cohens_d}", remove.bracket = FALSE,label.size=4)+
#     ylab("Pseudo Purity POST-treatment")+
#     xlab("Response")+
#     theme(legend.position = "right",
#           text = element_text(size = 20),
#           axis.text.x=element_blank(),
#           axis.title.x=element_blank())+
#     guides(fill=guide_legend(title="Response"))
# }
# 
# dev.off()

# png(sprintf("%s/%s",figure_dir,"IM_POST.png"),width = 480, height = 300)
# 
# wilcox_test_val <- compare_means(ImmuneScore ~ AX_BOR3, p.adjust.method = "BH", data=manifest_POST_filt,group.by = "TRTGRP")
# max_imm <- max(manifest_POST_filt$ImmuneScore,na.rm=T)
# wilcox_test_val$y.position <- rep(rep(max_imm,3)+seq(from=max_imm/6,to=(max_imm/6)*3,by=max_imm/6),3)
# cohens_d <- calculate_cohens_d(wilcox_test_val,c("TRTGRP"),"AX_BOR3",manifest_POST_filt)
# wilcox_test_val$cohens_d <- format(round(cohens_d$cohens_d, 2), nsmall = 2)
# wilcox_test_val <- wilcox_test_val %>% dplyr::filter(p.adj<0.05)
# if (nrow(wilcox_test_val)==0){
#   ggplot(manifest_POST_filt,aes(x=AX_BOR3,y=ImmuneScore))+
#     geom_boxplot(aes(fill=AX_BOR3))+geom_point()+facet_grid(.~TRTGRP)+
#     ylab("Imm. Score POST-treatment")+
#     xlab("Response")+
#     theme(legend.position = "bottom",
#           text = element_text(size = 20),
#           axis.text.x=element_blank(),
#           axis.title.x=element_blank())+
#     guides(fill=guide_legend(title="Response"))
# } else {
#   ggplot(manifest_POST_filt,aes(x=AX_BOR3,y=ImmuneScore))+
#     geom_boxplot(aes(fill=AX_BOR3))+geom_point()+facet_grid(.~TRTGRP)+
#     add_pvalue(wilcox_test_val, label = "p={p.adj},d={cohens_d}", remove.bracket = FALSE,label.size=4)+
#     ylab("Imm. Score POST-treatment")+
#     xlab("Response")+
#     theme(legend.position = "bottom",
#           text = element_text(size = 20),
#           axis.text.x=element_blank(),
#           axis.title.x=element_blank())+
#     guides(fill=guide_legend(title="Response"))
# }
# 
# dev.off()

# png(sprintf("%s/%s",figure_dir,"IM_POST_all.png"),width = 480/1.5, height = 300/1.25)
# 
# wilcox_test_val <- compare_means(ImmuneScore ~ AX_BOR3, p.adjust.method = "BH", data=manifest_POST_filt)
# max_imm <- max(manifest_POST_filt$ImmuneScore,na.rm=T)
# wilcox_test_val$y.position <- rep(max_imm,3)+seq(from=max_imm/6,to=(max_imm/6)*3,by=max_imm/6)
# cohens_d <- calculate_cohens_d(wilcox_test_val,c(),"AX_BOR3",manifest_POST_filt)
# wilcox_test_val$cohens_d <- format(round(cohens_d$cohens_d, 2), nsmall = 2)
# wilcox_test_val <- wilcox_test_val %>% dplyr::filter(p<0.05)
# if (nrow(wilcox_test_val)==0){
# ggplot(manifest_POST_filt,aes(x=AX_BOR3,y=ImmuneScore))+
#   geom_boxplot(aes(fill=AX_BOR3))+geom_point()+
#   ylab("Imm. Score POST-treatment")+
#   xlab("Response")+
#   theme(legend.position = "right",
#         text = element_text(size = 20),
#         axis.text.x=element_blank(),
#         axis.title.x=element_blank())+
#   guides(fill=guide_legend(title="Response"))
# } else {
#   ggplot(manifest_POST_filt,aes(x=AX_BOR3,y=ImmuneScore))+
#     geom_boxplot(aes(fill=AX_BOR3))+geom_point()+
#     add_pvalue(wilcox_test_val, label = "p={p.format},d={cohens_d}", remove.bracket = FALSE,label.size=4)+
#     ylab("Imm. Score POST-treatment")+
#     xlab("Response")+
#     theme(legend.position = "right",
#           text = element_text(size = 20),
#           axis.text.x=element_blank(),
#           axis.title.x=element_blank())+
#     guides(fill=guide_legend(title="Response"))
# }
# 
# dev.off()

# png(sprintf("%s/%s",figure_dir,"ST_POST.png"),width = 480, height = 300)
# 
# wilcox_test_val <- compare_means(StromalScore ~ AX_BOR3, p.adjust.method = "BH", data=manifest_POST_filt,group.by = "TRTGRP")
# max_strom <- max(manifest_POST_filt$StromalScore,na.rm=T)
# wilcox_test_val$y.position <- rep(rep(max_strom,3)+seq(from=max_strom/6,to=(max_strom/6)*3,by=max_strom/6),3)
# cohens_d <- calculate_cohens_d(wilcox_test_val,c("TRTGRP"),"AX_BOR3",manifest_POST_filt)
# wilcox_test_val$cohens_d <- format(round(cohens_d$cohens_d, 2), nsmall = 2)
# wilcox_test_val <- wilcox_test_val %>% dplyr::filter(p.adj<0.05)
# if (nrow(wilcox_test_val)==0){
#   ggplot(manifest_POST_filt,aes(x=AX_BOR3,y=StromalScore))+
#     geom_boxplot(aes(fill=AX_BOR3))+geom_point()+facet_grid(.~TRTGRP)+
#     ylab("Strom. Score POST-treatment")+
#     xlab("Response")+
#     theme(legend.position = "bottom",
#           text = element_text(size = 20),
#           axis.text.x=element_blank(),
#           axis.title.x=element_blank())+
#     guides(fill=guide_legend(title="Response"))
# } else {
#   ggplot(manifest_POST_filt,aes(x=AX_BOR3,y=StromalScore))+
#     geom_boxplot(aes(fill=AX_BOR3))+geom_point()+facet_grid(.~TRTGRP)+
#     add_pvalue(wilcox_test_val, label = "p={p.adj},d={cohens_d}", remove.bracket = FALSE,label.size=4)+
#     ylab("Strom. Score POST-treatment")+
#     xlab("Response")+
#     theme(legend.position = "bottom",
#           text = element_text(size = 20),
#           axis.text.x=element_blank(),
#           axis.title.x=element_blank())+
#     guides(fill=guide_legend(title="Response"))
# }
# 
# dev.off()

# png(sprintf("%s/%s",figure_dir,"ST_POST_all.png"),width = 480/1.5, height = 300/1.25)
# 
# wilcox_test_val <- compare_means(StromalScore ~ AX_BOR3, p.adjust.method = "BH", data=manifest_POST_filt)
# max_strom <- max(manifest_POST_filt$StromalScore,na.rm=T)
# wilcox_test_val$y.position <- rep(max_strom,3)+seq(from=max_strom/6,to=(max_strom/6)*3,by=max_strom/6)
# cohens_d <- calculate_cohens_d(wilcox_test_val,c(),"AX_BOR3",manifest_POST_filt)
# wilcox_test_val$cohens_d <- format(round(cohens_d$cohens_d, 2), nsmall = 2)
# wilcox_test_val <- wilcox_test_val %>% dplyr::filter(p<0.05)
# if (nrow(wilcox_test_val)==0){
# ggplot(manifest_POST_filt,aes(x=AX_BOR3,y=StromalScore))+
#   geom_boxplot(aes(fill=AX_BOR3))+geom_point()+
#   ylab("Strom. Score POST-treatment")+
#   xlab("Response")+
#   theme(legend.position = "right",
#         text = element_text(size = 20),
#         axis.text.x=element_blank(),
#         axis.title.x=element_blank())+
#   guides(fill=guide_legend(title="Response"))
# 
# } else {
#   ggplot(manifest_POST_filt,aes(x=AX_BOR3,y=StromalScore))+
#     geom_boxplot(aes(fill=AX_BOR3))+geom_point()+
#     add_pvalue(wilcox_test_val, label = "p={p.format},d={cohens_d}", remove.bracket = FALSE,label.size=4)+
#     ylab("Strom. Score POST-treatment")+
#     xlab("Response")+
#     theme(legend.position = "right",
#           text = element_text(size = 20),
#           axis.text.x=element_blank(),
#           axis.title.x=element_blank())+
#     guides(fill=guide_legend(title="Response"))
# }
# 
# dev.off()

# png(sprintf("%s/%s",figure_dir,"SA_mean_PP_POST_all.png"),width = 480, height = 300)
# 
# ggplot(manifest_POST,aes(x=pseudo_purity,y=SA_mean))+geom_point()+
#   stat_cor(method = "kendall",label.x.npc="middle",label.y.npc="top",size = 5)+
#   geom_smooth(method="lm",se=F)+
#   theme(text = element_text(size = 20))
# 
# dev.off()

# png(sprintf("%s/%s",figure_dir,"SA_mean_PP_POST.png"),width = 480, height = 300)
# 
# ggplot(manifest_POST,aes(x=pseudo_purity,y=SA_mean))+geom_point()+
#   stat_cor(method = "kendall",label.x.npc="left",label.y.npc="top",size = 4)+
#   geom_smooth(method="lm",se=F)+facet_grid(AX_BOR3~TRTGRP)+
#   theme(text = element_text(size = 20))
# 
# dev.off()

png(sprintf("%s/%s",figure_dir,"Fig5A.png"),width = 480, height = 300)

ggplot(manifest_POST,aes(x=pseudo_purity,y=SA_mean))+geom_point()+
  stat_cor(method = "kendall",label.x.npc="left",label.y.npc="top",size = 5)+
  geom_smooth(method="lm",se=F)+facet_grid(AX_BOR3~.)+
  theme(text = element_text(size = 20))

dev.off()

# png(sprintf("%s/%s",figure_dir,"SA_mean_IM_POST_all.png"),width = 480, height = 300)
# 
# ggplot(manifest_POST,aes(x=ImmuneScore,y=SA_mean))+geom_point()+
#   stat_cor(method = "kendall",label.x.npc="middle",label.y.npc="top",size = 5)+
#   geom_smooth(method="lm",se=F)+
#   theme(text = element_text(size = 20))
# 
# dev.off()

# png(sprintf("%s/%s",figure_dir,"SA_mean_IM_POST.png"),width = 480, height = 300)
# 
# ggplot(manifest_POST,aes(x=ImmuneScore,y=SA_mean))+geom_point()+
#   stat_cor(method = "kendall",label.x.npc="left",label.y.npc="top",size = 4)+
#   geom_smooth(method="lm",se=F)+facet_grid(AX_BOR3~TRTGRP)+
#   theme(text = element_text(size = 20))
# 
# dev.off()

# png(sprintf("%s/%s",figure_dir,"SA_mean_IM_POST_resp.png"),width = 480, height = 300)
# 
# ggplot(manifest_POST,aes(x=ImmuneScore,y=SA_mean))+geom_point()+
#   stat_cor(method = "kendall",label.x.npc="left",label.y.npc="top",size = 5)+
#   geom_smooth(method="lm",se=F)+facet_grid(AX_BOR3~.)+
#   theme(text = element_text(size = 20))
# 
# dev.off()

# png(sprintf("%s/%s",figure_dir,"SA_mean_ST_POST_all.png"),width = 480, height = 300)
# 
# ggplot(manifest_POST,aes(x=StromalScore,y=SA_mean))+geom_point()+
#   stat_cor(method = "kendall",label.x.npc="middle",label.y.npc="top",size = 5)+
#   geom_smooth(method="lm",se=F)+
#   theme(text = element_text(size = 20))
# 
# dev.off()

# png(sprintf("%s/%s",figure_dir,"SA_mean_ST_POST.png"),width = 480, height = 300)
# 
# ggplot(manifest_POST,aes(x=StromalScore,y=SA_mean))+geom_point()+
#   stat_cor(method = "kendall",label.x.npc="left",label.y.npc="top",size = 5)+
#   geom_smooth(method="lm",se=F)+facet_grid(AX_BOR3~TRTGRP)+
#   theme(text = element_text(size = 20))
# 
# dev.off()

# png(sprintf("%s/%s",figure_dir,"SA_mean_ST_POST_resp.png"),width = 480, height = 300)
# 
# ggplot(manifest_POST,aes(x=StromalScore,y=SA_mean))+geom_point()+
#   stat_cor(method = "kendall",label.x.npc="left",label.y.npc="top",size = 5)+
#   geom_smooth(method="lm",se=F)+facet_grid(AX_BOR3~.)+
#   theme(text = element_text(size = 20))
# 
# dev.off()

# png(sprintf("%s/%s",figure_dir,"PP_TP_POST_all.png"),width = 480, height = 300)
# 
# ggplot(manifest_POST,aes(x=tumor_purity,y=pseudo_purity))+geom_point()+
#   stat_cor(method = "kendall",label.x.npc="middle",label.y.npc="top",size = 5)+
#   geom_smooth(method="lm",se=F)+
#   theme(text = element_text(size = 20))
# 
# dev.off()

```
# Splicing antigenicity per event and sample

```{r}

comparison_directory <- "./input_data/create_comparisons_out_cp"
kmers_file <- sprintf("%s/kmers_specific_baseline_vs_post_treatment.rds",comparison_directory)
splice_dat_file <- sprintf("%s/splice_dat_baseline_vs_post_treatment.rds",comparison_directory)
junc_expr_file <- sprintf("%s/junc_expr_combined_vst.rds",comparison_directory)

kmers <- readRDS(kmers_file)
splice_dat <- readRDS(splice_dat_file)
junc_expr <- readRDS(junc_expr_file)
colnames(junc_expr) <- str_remove(colnames(junc_expr),".junc")

kmer_juncs <- kmers$juncs
kmer_samples <- colnames(kmers)[seq(3,ncol(kmers))]

junc_expr_kmer_matched <- junc_expr[kmer_juncs,kmer_samples]

convert_to_numeric <- function(df){
  for (col in seq(ncol(df))){
    df[,col]<-as.numeric(df[,col])
  }
  return(df)
}
kmer_numeric <- convert_to_numeric(kmers[,seq(3,ncol(kmers))])
rownames(kmer_numeric)<-seq(nrow(kmer_numeric))
junc_numeric <- convert_to_numeric(junc_expr_kmer_matched)
rownames(junc_numeric)<-seq(nrow(junc_numeric))
SA_per_junction <- kmer_numeric*junc_numeric
SA_per_junction_norm <- SA_per_junction
for (col_name in colnames(SA_per_junction)){
  purity <- manifest$pseudo_purity[manifest$Sample==col_name]
  SA_per_junction_norm[,col_name] <- SA_per_junction[,col_name,drop=T]/purity
}
SA_per_junction_norm$rows <- kmers$rows
SA_per_junction_norm$juncs <- kmers$juncs
SA_per_junction_norm$gene <- splice_dat$gene


top_genes <- unique(unname(unlist(lapply(seq(ncol(splicing_antigenicity)),function(col_val){
  genes <- rownames(splicing_antigenicity)
  SA_vec <- splicing_antigenicity[,col_val,drop=T]
  names(SA_vec)<-genes
  top_genes <- names(sort(SA_vec))[seq(20)]
  return(top_genes)
}))))

SA_per_junction_filt <- SA_per_junction_norm %>% dplyr::filter(gene %in% top_genes)
splice_dat_filt <- splice_dat %>% dplyr::filter(gene %in% top_genes)

```

## Per junction splicing antigenicity analysis

```{r}

CRPR_POST <- manifest$Sample[manifest$AX_TIMETEMP=="POST" & manifest$AX_BOR3=="CRPR" & !manifest$AX_BOR=="NE"]
SD_POST <- manifest$Sample[manifest$AX_TIMETEMP=="POST" & manifest$AX_BOR3=="SD" & !manifest$AX_BOR=="NE"]
PD_POST <- manifest$Sample[manifest$AX_TIMETEMP=="POST" & manifest$AX_BOR3=="PD" & !manifest$AX_BOR=="NE"]
PRE_samples <- manifest$Sample[manifest$AX_PriorIPI==0 & manifest$AX_TIMETEMP=="PRE"]

CRPR_POST_SA_mean <- data.frame(mean=apply(SA_per_junction_filt[,CRPR_POST],1,mean),
                                  juncs=SA_per_junction_filt$juncs,
                                  rows=SA_per_junction_filt$rows,
                                  gene=SA_per_junction_filt$gene,
                                  response=rep("CRPR",nrow(SA_per_junction_filt)))
SD_POST_SA_mean <- data.frame(mean=apply(SA_per_junction_filt[,SD_POST],1,mean),
                                  juncs=SA_per_junction_filt$juncs,
                                  rows=SA_per_junction_filt$rows,
                                  gene=SA_per_junction_filt$gene,
                                  response=rep("SD",nrow(SA_per_junction_filt)))
PD_POST_SA_mean <- data.frame(mean=apply(SA_per_junction_filt[,PD_POST],1,mean),
                                  juncs=SA_per_junction_filt$juncs,
                                  rows=SA_per_junction_filt$rows,
                                  gene=SA_per_junction_filt$gene,
                                  response=rep("PD",nrow(SA_per_junction_filt)))
PRE_SA_mean <- data.frame(mean=apply(SA_per_junction_filt[,PRE_samples],1,mean),
                                  juncs=SA_per_junction_filt$juncs,
                                  rows=SA_per_junction_filt$rows,
                                  gene=SA_per_junction_filt$gene,
                                  response=rep("baseline",nrow(SA_per_junction_filt)))

means <- rbind(CRPR_POST_SA_mean,
                 SD_POST_SA_mean,
                 PD_POST_SA_mean,
                 PRE_SA_mean)
means_wilcox <- compare_means(mean~response,data=means,paired=T)
cohens_d <- calculate_cohens_d(means_wilcox,c(),"response",means)
means_wilcox$cohens_d <- format(round(cohens_d$cohens_d, 2), nsmall = 2)
max_SA <- max(means$mean)
means_wilcox$y.position <- log10(rep(max_SA,6))+c(1,2,3,4,5,6)*0.5


CRPR_POST_SA_median <- data.frame(median=apply(SA_per_junction_filt[,CRPR_POST],1,median),
                                  juncs=SA_per_junction_filt$juncs,
                                  rows=SA_per_junction_filt$rows,
                                  gene=SA_per_junction_filt$gene,
                                  response=rep("CRPR",nrow(SA_per_junction_filt)))
SD_POST_SA_median <- data.frame(median=apply(SA_per_junction_filt[,SD_POST],1,median),
                                  juncs=SA_per_junction_filt$juncs,
                                  rows=SA_per_junction_filt$rows,
                                  gene=SA_per_junction_filt$gene,
                                  response=rep("SD",nrow(SA_per_junction_filt)))
PD_POST_SA_median <- data.frame(median=apply(SA_per_junction_filt[,PD_POST],1,median),
                                  juncs=SA_per_junction_filt$juncs,
                                  rows=SA_per_junction_filt$rows,
                                  gene=SA_per_junction_filt$gene,
                                  response=rep("PD",nrow(SA_per_junction_filt)))
PRE_SA_median <- data.frame(median=apply(SA_per_junction_filt[,PRE_samples],1,median),
                                  juncs=SA_per_junction_filt$juncs,
                                  rows=SA_per_junction_filt$rows,
                                  gene=SA_per_junction_filt$gene,
                                  response=rep("baseline",nrow(SA_per_junction_filt)))

medians <- rbind(CRPR_POST_SA_median,
                 SD_POST_SA_median,
                 PD_POST_SA_median,
                 PRE_SA_median)

medians_wilcox <- compare_means(median~response,data=medians,paired=T)
cohens_d <- calculate_cohens_d(medians_wilcox,c(),"response",medians)
medians_wilcox$cohens_d <- format(round(cohens_d$cohens_d, 2), nsmall = 2)


```

## Plotting pairwise comparisons

```{r}

figure_dir <- "./figures"

png(sprintf("%s/%s",figure_dir,"Fig6A.png"),width = 480/1.5, height = 300)

medians_sorted <- medians
medians_sorted <- medians_sorted %>% dplyr::filter(response %in% c("CRPR","baseline"))
medians_wilcox_filt <- medians_wilcox %>% dplyr::filter(group1 == "CRPR" & group2 == "baseline")

medians_sorted$response <- factor(medians_sorted$response,levels=c("CRPR","baseline"))
max_median <- max(medians$median)
ggplot(medians_sorted,aes(x=response,y=log10(median)))+
  geom_boxplot(aes(fill=response))+geom_line(aes(group=juncs),alpha=0.05)+
  add_pvalue(medians_wilcox_filt, label = "{p.signif},d={cohens_d}", remove.bracket = FALSE,label.size=6,y.position=log10(max_median)+1)+
  scale_color_manual(values = c("CRPR" = "#F8766D",
                              "SD"="#00BA38",
                              "PD"="#619CFF",
                              "baseline"="wheat"),
                     aesthetics="fill")+
    theme(legend.position = "bottom",
        text = element_text(size = 20),
        axis.text.x=element_blank(),
        axis.title.x=element_blank())+
  guides(fill=guide_legend(title="Response"))+
  ylab("log10(Junc. SA norm)")+
  ylim(c(0,log10(max_median)+1.5))+
  geom_hline(yintercept = 3)

dev.off()

png(sprintf("%s/%s",figure_dir,"Fig6B.png"),width = 480/1.5, height = 300)

medians_sorted <- medians
medians_sorted <- medians_sorted %>% dplyr::filter(response %in% c("SD","baseline"))
medians_wilcox_filt <- medians_wilcox %>% dplyr::filter(group1 == "SD" & group2 == "baseline")
medians_sorted$response <- factor(medians_sorted$response,levels=c("SD","baseline"))
max_median <- max(medians$median)
ggplot(medians_sorted,aes(x=response,y=log10(median)))+
  geom_boxplot(aes(fill=response))+geom_line(aes(group=juncs),alpha=0.05)+
  add_pvalue(medians_wilcox_filt, label = "{p.signif},d={cohens_d}", remove.bracket = FALSE,label.size=6,y.position=log10(max_median)+1)+
  scale_color_manual(values = c("CRPR" = "#F8766D",
                              "SD"="#00BA38",
                              "PD"="#619CFF",
                              "baseline"="wheat"),
                     aesthetics="fill")+
    theme(legend.position = "bottom",
        text = element_text(size = 20),
        axis.text.x=element_blank(),
        axis.title.x=element_blank())+
  guides(fill=guide_legend(title="Response"))+
  ylab("log10(Junc. SA norm)")+
  ylim(c(0,log10(max_median)+1.5))+
  geom_hline(yintercept = 3)

dev.off()

png(sprintf("%s/%s",figure_dir,"Fig6C.png"),width = 480/1.5, height = 300)

medians_sorted <- medians
medians_sorted <- medians_sorted %>% dplyr::filter(response %in% c("PD","baseline"))
medians_wilcox_filt <- medians_wilcox %>% dplyr::filter(group1 == "PD" & group2 == "baseline")
medians_sorted$response <- factor(medians_sorted$response,levels=c("PD","baseline"))
max_median <- max(medians$median)
ggplot(medians_sorted,aes(x=response,y=log10(median)))+
  geom_boxplot(aes(fill=response))+geom_line(aes(group=juncs),alpha=0.05)+
  add_pvalue(medians_wilcox_filt, label = "{p.signif},d={cohens_d}", remove.bracket = FALSE,label.size=6,y.position=log10(max_median)+1)+
  scale_color_manual(values = c("CRPR" = "#F8766D",
                              "SD"="#00BA38",
                              "PD"="#619CFF",
                              "baseline"="wheat"),
                     aesthetics="fill")+
    theme(legend.position = "bottom",
        text = element_text(size = 20),
        axis.text.x=element_blank(),
        axis.title.x=element_blank())+
  guides(fill=guide_legend(title="Response"))+
  ylab("log10(Junc. SA norm)")+
  ylim(c(0,log10(max_median)+1.5))+
  geom_hline(yintercept = 3)

dev.off()

```

```{r}

print(sessionInfo())

```

